% Copyright (c) 2014-2025 Yegor Bugayenko
%
% Permission is hereby granted, free of charge, to any person obtaining a copy
% of this software and associated documentation files (the 'Software'), to deal
% in the Software without restriction, including without limitation the rights
% to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
% copies of the Software, and to permit persons to whom the Software is
% furnished to do so. The Software doesn't include files with .md extension.
% That files you are not allowed to copy, distribute, modify, publish, or sell.
%
% THE SOFTWARE IS PROVIDED 'AS IS', WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
% IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
% FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
% AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
% LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
% OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
% SOFTWARE.

\documentclass{./yb}
\newcommand\first[2]{\vspace{1em}{\setlength{\parindent}{0pt}\bfseries\LARGE #1}\textsc{#2}}

\begin{document}

{\bfseries\scshape Teaching Statement}\newline
Yegor Bugayenko\newline
\today

\vspace*{12pt}

\first{I}{ don't} believe in teaching. I believe in learning, which
just happens when someone has motivation and ability.
Well, I think I'm wrong about the ability---just motivation!
Once there is a strong \textbf{motive}, the student is in the
driver's seat, while teachers, mentors, books, and video
courses are mere assistants, at best.

A motive is the derivative of the answer, which so many students
struggle to find for their most important simple
question: \emph{``What's in it for me?''}
Indeed, why do they need to study
the relational theory, distributed transaction processing,
Boolean algebra, or $\lambda$-calculus?

``Because you need a diploma,'' is the easiest response some
teachers resort to. I remember myself being a student 25 years ago---such
an answer was not at all convincing for me. No matter how harsh
the exams were, how strict the discipline was, and how much I paid
for the study---I needed something else. And I had it.

Obviously, every student has their own motive.
Quite often it's a very unique one.
Some of them dream about their own
career in science and the answer would be ``That's how you
become a professor.'' Some of them imagine themselves as unicorn startup
founders---their answer would be ``It will make you rich.''
Some of them don't even know who and where they want to be in a few
decades---their motive is yet to be discovered.
No matter who is in front of me, my first and most
important goal as their teacher is to understand what they dream about
and then explain how learning the subject may help them get there.

\first{O}{nce} they know where they are going and why, they need
to know how to measure their progress.
They need some \textbf{metrics} of success.
They want to know whether the dream is getting closer or they are stuck.

``Just get a passing grade!'' is what some teachers use as a
scale of measurement. I see it differently. Honestly, I don't see myself
as the most competent in the subjects I teach. In each of
them---be it object-oriented programming, software architecture, DevOps, or project management---there
are world-class experts sitting on program committees of international conferences and academic journals.
They, I believe, must be the judges of my students.
They are the best providers of knowledge and expertise the market has to offer.

My job, as a teacher, is to introduce my students to the market, which will
judge them, both while they study and then later, when they graduate and work
professionally. I have to explain and
demonstrate to them how to submit their work to conferences, how to make
their code open source, how to present their work to funding committees,
how to publish their first blog posts, and so on.

I believe that the size of my classroom is too small for most talented students.
I have to introduce them to the rest of the world, which some of them will conquer, eventually.

\first{W}{hen} they have a reason to learn and the market, not me, appraises
them, I become their \textbf{peer}. Just like them, I try to publish
\href{https://www.yegor256.com/papers.html}{my work} at conferences,
I try to make \href{https://github.com/yegor256}{my software}
popular on GitHub, I try to learn new programming
languages, and try to get funding for my research projects.
We read the same books, watch the same YouTube videos, and attend the same webinars.
The only difference between us: they are much younger. They
think faster, they take risks more easily, and they love JavaScript.

Almost always, they need me as a walking addendum to Wikipedia:
when something is unclear, they expect me to explain. When I do it,
I see myself not as their teacher, but as just a slightly more experienced colleague.
I don't teach them, I share what I know. Just like I share
\href{https://stackoverflow.com/users/187141/yegor256}{on Stack Overflow}
and on \href{https://www.yegor256.com}{my blog}.

That's how we learn together.

\end{document}
